---
title: "Homework 3"
author: "Yan Song"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(hexbin)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "80%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Due date

Due: October 10 at 10:00pm. 

### Points

| Problem         | Points    |
|:--------------- |:--------- |
| Problem 0       | 20        |
| Problem 1       | --        |
| Problem 2       | 40        |
| Problem 3       | 40        |
| Optional survey | No points |


### Problem 0

This "problem" focuses on structure of your submission, especially the use git and GitHub for reproducibility, R Projects to organize your work, R Markdown to write reproducible reports, relative paths to load data from local files, and reasonable naming structures for your files. 


### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and ... columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```

********

Let's make a plot

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

********

Let's make a table!!

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```

********

Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```
### Problem 2
 
```{r laod and tidy}
accel_data = 
  read_csv(file = "./data/accel_data.csv") %>% 
  pivot_longer(
    activity.1:activity.1440,
    names_to = "minute",
    names_prefix = "activity.",
    values_to = "activity"
  ) %>% 
  mutate(
    minute = as.numeric(minute),
    day_1 = day, 
    day_1 = recode(day, "Monday" = "1" ,"Tuesday" =  "2", "Wednesday" = "3", "Thursday" = "4", "Friday" = "5", "Saturday" = "6", "Sunday" = "7"),
    weekday_type = ifelse(day_1 > 5, "weekend", "weekday"),
    day = factor(day, order = TRUE, levels = c("Monday","Tuesday","Wednesday", "Thursday" , "Friday", "Saturday","Sunday"))
  ) %>% 
  arrange(day) %>% 
  select(-day_1)
```

There is `r nrow(accel_data)` rows and `r ncol(accel_data)` columns in this dataset. 
The variables contains :
*week: from 1 to 5
*day_id: from 1 to 35
*day: Monday to Sunday
*weekday_type:(weekday or weekend):
*minute(the minute of the day):
*activity: the activity counts for each minute

There is `r nrow(accel_data)` observation.

********

Create a readable table showing total activity for each day.
```{r}
total_activity_a_day = 
  accel_data %>% 
  group_by(week,day) %>% 
  summarize(
    total_each_day = sum(activity)
  )


table_for_total = 
  total_activity_a_day %>% 
  pivot_wider(
    names_from = "day",
    values_from = "total_each_day"
  )

knitr::kable(table_for_total, digits = 2)

```


The trends are not apparent.

**********

Draw a plot to see the trends.
```{r}
total_activity_a_day %>% 
  ungroup(week,day) %>% 
  mutate(day_number = row_number()) %>% 
  ggplot(aes(x = day_number, y = total_each_day, color = week)) + 
  geom_point() + 
  geom_line() + 
  scale_y_continuous(
  breaks = c(100000,200000,300000,400000,500000,600000,700000,800000), 
  labels = c("100000","200000","300000","400000","500000","600000","700000", "800000")
    ) + 
  labs(
    x = "Day",
    y = "Total activity",
    caption = "Data from the Advanced Cardiac Care Center of Columbia University Medical Center"
  )

```

Now we could see the trends of total activity for each day.

********

Draw a single-panel plot that shows the 24-hour activity time courses for each day.

```{r}
accel_data %>% 
  mutate(
    hour = ceiling(minute/60)
  ) %>% 
  group_by(day_id, hour) %>% 
  mutate(mean_act = mean(activity)) %>% 
  ggplot(aes(x = hour, y = mean_act, color = day, group = day_id)) +
  geom_point()+
  geom_line() +
  labs(
    x = "Hour of each day",
    y = "Activity",
    caption = "Data from the Advanced Cardiac Care Center of Columbia University Medical Center" )
```

We could see that this man's activity counts was high in 9pm on Friday and was inactive during 0am to 8am everyday.

### Problem 3

```{r}
library(p8105.datasets)
data("ny_noaa")
```

```{r}
weather_df= ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date,into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.abb[month],
    tmin = as.numeric(tmin),
    tmax = as.numeric(tmax),
    tmax = tmax / 10,
    tmin = tmin / 10,
    prcp = prcp / 10
)

```

```{r}
snow = weather_df$snow 
snow_mode= names(table(snow))[table(snow)==max(table(snow))]  
```
For snowfall, what are the most commonly observed values?

********
Make a two-panel plots showing the average max temperature in January and July in each station across years.

```{r}
weather_df %>%
filter(month %in% c("Jan", "Jul")) %>%
group_by(id, year, month) %>%
summarize(avtmaxd = mean(tmax, na.rm = TRUE), .groups = 'drop') %>%
ggplot(aes(x = year, y = avtmaxd, group = id)) +
geom_line() +
facet_grid(month ~.)+
labs(title = "Mean Average Temperature for January and July Across Stations and Years", x = "Year", y = "Average Maximum Temperature (C)") +
theme(axis.title = element_text(size = 8),
      plot.title = element_text(size = 10))
```

********
Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}

tmax_min_p = weather_df %>%
  drop_na(tmin,tmax)%>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()+
  theme (legend.position = "right",
    legend.title = element_text(size=10)
  )

snowfall_df = weather_df %>%
  filter (0 < snow, snow < 100) %>%
  group_by(year) %>%
  select(year, snow) %>% 
  mutate(year = as.character(year)) %>% 
  ggplot(aes(x= year, y= snow)) +
  geom_boxplot()+
  theme(axis.text.x = element_text (angle = 90, vjust = 0.5, hjust = 1))
tmax_min_p / snowfall_df 

```

