---
title: "Homework 3"
author: "Yan Song"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(hexbin)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "80%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Due date

Due: October 10 at 10:00pm. 

### Points

| Problem         | Points    |
|:--------------- |:--------- |
| Problem 0       | 20        |
| Problem 1       | --        |
| Problem 2       | 40        |
| Problem 3       | 40        |
| Optional survey | No points |


### Problem 0

This "problem" focuses on structure of your submission, especially the use git and GitHub for reproducibility, R Projects to organize your work, R Markdown to write reproducible reports, relative paths to load data from local files, and reasonable naming structures for your files. 


### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

********

How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```

Let's make a plot

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Let's make a table!!

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```
### Problem 2

Read and clean accelerometers dataset. 
```{r laod and tidy}
accel_data = 
  read_csv(file = "./data/accel_data.csv") %>% 
  pivot_longer(
    activity.1:activity.1440,
    names_to = "minute",
    names_prefix = "activity.",
    values_to = "activity"
  ) %>% 
  mutate(
    minute = as.numeric(minute),
    day_1 = day, 
    day_1 = recode(day, "Monday" = "1" ,"Tuesday" =  "2", "Wednesday" = "3", "Thursday" = "4", "Friday" = "5", "Saturday" = "6", "Sunday" = "7"),
    weekday_type = ifelse(day_1 > 5, "weekend", "weekday"),
    day = factor(day, order = TRUE, levels = c("Monday","Tuesday","Wednesday", "Thursday" , "Friday", "Saturday","Sunday"))
  ) %>% 
  arrange(day) %>% 
  select(-day_1)
```

There are `r nrow(accel_data)` rows and `r ncol(accel_data)` columns in this dataset while `r nrow(accel_data)` observations in total.

The dataset includes the following variables:

* week: from 1 to 5
* day_id: from 1 to 35
* day: Monday to Sunday
* weekday_type: weekday or weekend
* minute: the minute of the day
* activity: the activity counts for each minute

********

Create a readable table showing total activity for each day.
```{r}
total_activity_a_day = 
  accel_data %>% 
  group_by(week,day) %>% 
  summarize(
    total_each_day = sum(activity)
  )


table_for_total = 
  total_activity_a_day %>% 
  pivot_wider(
    names_from = "day",
    values_from = "total_each_day"
  )

knitr::kable(table_for_total, digits = 2)

```

The trends are not apparent.

**********

Draw a plot to see the trends.
```{r}
total_activity_a_day %>% 
  ungroup(week,day) %>% 
  mutate(day_number = row_number()) %>% 
  ggplot(aes(x = day_number, y = total_each_day, color = week)) + 
  geom_point() + 
  geom_line() + 
  scale_y_continuous(
  breaks = c(100000,200000,300000,400000,500000,600000,700000,800000), 
  labels = c("100000","200000","300000","400000","500000","600000","700000", "800000")
    ) + 
  labs(
    x = "Day",
    y = "Total Activity",
    caption = "Data from the Advanced Cardiac Care Center of Columbia University Medical Center"
  )+
  theme (plot.caption = element_text(hjust = 1, size = 6, face = "italic"))

```

Now we could see the trends of total activity for each day.

********

Draw a single-panel plot that shows the 24-hour activity time courses for each day.

```{r}
accel_data %>% 
  mutate(
    hour = ceiling(minute/60)
  ) %>% 
  group_by(day_id, hour) %>% 
  mutate(mean_act = mean(activity)) %>% 
  ggplot(aes(x = hour, y = mean_act, color = day, group = day_id)) +
  geom_point()+
  geom_line() +
  labs(
    x = "Hour of Each Day",
    y = "Activity",
    caption = "Data from the Advanced Cardiac Care Center of Columbia University Medical Center" )+
   theme(plot.caption = element_text(hjust = 1, size = 6, face = "italic"))
```

We could see that this man's activity counts was high in 9pm on Friday and was inactive during 0am to 8am everyday.

### Problem 3

Load NY NOAA data.
```{r}
library(p8105.datasets)
data("ny_noaa")
```

Clean the data and create separate variables for year, month, and day. 
```{r}
weather_df= ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date,into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.abb[month],
    tmin = as.numeric(tmin),
    tmax = as.numeric(tmax),
    tmax = tmax / 10,
    tmin = tmin / 10,
    prcp = prcp / 10)

```
NY NOAA data contains core variables for all New York state weather station form January 1, 1981 through December 31, 2010. There are `r nrow(weather_df)` rows and `r ncol(weather_df)` columns in this dataset while `r nrow(weather_df)` observations in total. 

The dataset contains variables:

* id: Weather station ID 
* date: Date of observation
* prcp: Precipitation (tenths of mm)
* snow: Snowfall (mm)
* snwd: Snow depth (mm)
* tmax: Maximum temperature (tenths of degrees C)
* tmin: Minimum temperature (tenths of degrees C)

********

Next, check the most commonly observed values for snowfall.
```{r}
snow = weather_df$snow 
snow_mode= names(table(snow))[table(snow) == max(table(snow))]  
```
For snowfall, the most commonly observed values is `r snow_mode` mm.

********

Make a two-panel plots showing the average max temperature in January and July in each station across years.

```{r}
weather_df %>%
filter(month %in% c("Jan", "Jul")) %>%
group_by(id, year, month) %>%
summarize(avtmaxd = mean(tmax, na.rm = TRUE), .groups = 'drop') %>%
ggplot(aes(x = year, y = avtmaxd, group = id)) +
geom_line()+
geom_point()+
facet_grid(month ~.)+
labs(title = "Mean Average Temperature for January and July Across Stations and Years", x = "Year", y = "Average Maximum Temperature (C)") +
theme(axis.title = element_text(size = 8),
      plot.title = element_text(size = 10))
```

The plot shows increasing trends in the average maximum temperature in January and in July across the years, which could be an indicator of global warming. Also, there are several outliers in the January plot in 1982, 2005 that show much lower average maximum temperature to the other years while some outliers in 2004 has much higher average maximum temperature than the other years. In the plot of July, there are several outliers in the year of 1988, 2004 and 2007 show lower average maximum temperature to other years.

********

Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}

tmax_min_p = weather_df %>%
  drop_na(tmin,tmax)%>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()+
  labs(x = "Minimum Temperature (tenths of degrees C)", y = " Maximum Temperature (tenths of degrees C)")+
  theme (legend.position = "right",
    legend.title = element_text(size = 10),
    axis.title = element_text(size = 5)
  )

snowfall_df = weather_df %>%
  filter (0 < snow, snow < 100) %>%
  group_by(year) %>%
  select(year, snow) %>% 
  mutate(year = as.character(year)) %>% 
  ggplot(aes(x= year, y= snow)) +
  geom_boxplot()+
  labs(x = "Year", y = "Snowfall (mm)") +
  theme(axis.text.x = element_text (angle = 90, vjust = 0.5, hjust = 1),
    axis.title = element_text(size = 8))
      
tmax_min_p / snowfall_df 

```

Description:

* The average precipitation is `r round(mean(pull(weather_df, prcp), na.rm = TRUE), digits = 2)` tenths of mm with the standard deviation of `r round(sd(pull(weather_df, prcp), na.rm = TRUE), digits = 2)` tenths of mm. The maximum and minimum precipitation during the period is `r max(pull(weather_df,prcp), na.rm = TRUE)` tenths of mm and `r min(pull(weather_df,prcp), na.rm = TRUE)` tenths of mm respectively. 

* The average snowfall is `r round(mean(pull(weather_df, snow), na.rm = TRUE), digits = 2)` mm with the standard deviation of `r round(sd(pull(weather_df, snow), na.rm = TRUE), digits = 2)` mm. The maximum and minimum snowfall during the period is `r max(pull(weather_df, snow), na.rm = TRUE)` mm and `r min(pull(weather_df, snow), na.rm = TRUE)` mm respectively. 
 
* The average maximum temperature is `r round(mean(pull(weather_df, tmax), na.rm = TRUE), digits = 2)` tenths of degrees C with the standard deviation of `r round(sd(pull(weather_df, tmax), na.rm = TRUE), digits = 2)` tenths of degrees C. The range of maximum temperature during the period is between `r max(pull(weather_df, tmax), na.rm = TRUE)` tenths of degrees C and `r min(pull(weather_df, tmax), na.rm = TRUE)` tenths of degrees C. 

* The average minimum temperature is `r round(mean(pull(weather_df, tmin), na.rm = TRUE), digits = 2)` tenths of degrees C with the standard deviation of `r round(sd(pull(weather_df, tmin), na.rm = TRUE), digits = 2)` tenths of degrees C. The range of minimum temperature during the period is between  `r max(pull(weather_df, tmin), na.rm = TRUE)` tenths of degrees C and `r min(pull(weather_df, tmin), na.rm = TRUE)` tenths of degrees C. 

* There are `r sum(is.na(weather_df$tmin))` , `r sum(is.na(weather_df$tmax))` and `r sum(is.na(weather_df$snow))` missing values in the variables of minimum and maximum temperature, and snowfall respectively. The large number of missing values in the above variables might cause potential issues when drawing plots (e.g. the warning message pop up when drawing average max temperature in January and July in each station across years). 

